---
title: "An overview of my data science experience focusing on improvements to processes, relating this to any background knowledge of Geology"
output: 
  ioslides_presentation:
    widescreen: true
    smaller: true
    css: style.css
author: "Liubomyr Gavryliv"
date: Dec 17, 2020
autosize: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction | A word about myself

- Geological background
  - 2019: PhD in Geology ("Cenozoic mafic dykes of the Argentine Islands (West Antarctica)")
  - 2015: MSc in Geochemistry, Mineralogy and Petrography
  - 2013: BSc in geoinformatics

- Data science related experience: 
  - Since 2017: Freelancer on <a href="upwork.com">Upwork</a>.
  - Since 2018: Data manager/web developer at <a href="https://e-rocks.com">e-Rocks ltd</a>.
  
## Relational mineralogical data model | A cure for geoscientific community?

<h1 class="centered">Model characteristics</h1></br>

  - **Simple** – to be understood by geoscientists
  - **Customizable** – to be responsive to the needs of the researcher
  - **Relational and nested** – the data is organized into separate domains with intra- and cross-domain relations
  - **FAIR** – meet principles of findability, accessibility, interoperability, and reusability
  - **Fast** - find minerals on-the-fly

<h1 class="centered">Processes behind the scenes</h1></br>

  - **Data** (gather, ETL and enrich, validate and migrate to the DB)
  - **Back-end** service to speak nicely with the DB
  - **Front-end** layer to speak nicely with API
  - Model approbation in performing the **data analysis**

## Data

The data warehouse resides in a *postgreSQL* database at **AWS RDS**.

<h1 class="centered">Data gathering</h1></br>

The data comes in raw format from at least 3 different sources:

  - the existing mineralogical resources (eg <a href="http://rruff.geo.arizona.edu/AMS/amcsd.php">AMCSD</a>, 
  <a href="rruff.info/evolution">MED</a>, <a href="https://rruff.info/ima/">IMA list</a> etc)
  - web scraping open-source resources using *node.js* and *puppeteer.js*
  - hard-copy/online publications (<a href="http://www.minsocam.org/msa/ammin/ammineral.html">American Mineralogist</a>, <a href="https://pubs.geoscienceworld.org/canmin">Canadian Mineralogist</a> etc)
  
## Data ETL and enrich | <a href="https://github.com/liubomyrgavryliv/database-export">https://github.com/liubomyrgavryliv/database-export</a>

Initially, the data is loaded to *RStudio* and tested through next set of rules to make sure it meets the requirements of data warehouse:

  - join with **master table** of mineral species through mineral names key 
    - create 'anti-join' subset for double check
  - check duplicates
  - aggregate data - check unique values
  - clean data
    - remove trailing spaces
  - normalize data
    - transposing or pivoting (turning multiple columns into multiple rows or vice versa)
    - splitting a column into multiple columns (split strings into a list => explode a list into individual values in different columns)
  - join **master tables**, replace values with IDs from master tables
  
## Data validation, update/insert/delete on the DB side | <a href="https://github.com/liubomyrgavryliv/database-dev">https://github.com/liubomyrgavryliv/database-dev</a>

*Python* scripts to keep the DB synched and upload new batches of validated data. 
  
  - functions to generate import tables
  - *validator* class
    - cross-check new batch of data with **master tables** in the DB
    - record validation errors
    - create "pretty" validation report which is sent to a mineralogist for a review
  - *updater* class
    - decide which event should occur depending on table type (if master -> UPDATE/INSERT/DELETE, if many-to-many -> INSERT and/or DELETE)
    - record update events

## Example of data ETL to DB

## Mineralogical Nickel-Strunz classification | d3.js 

<iframe width="180" height="400" src="https://liubomyrgavryliv.github.io/d3-NS-sunburst-webpack/" scrolling="yes" frameborder="0" allowfullscreen></iframe>

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

